---
title: "June14_Kazu"
author: "Kazu"
date: "6/8/2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

```
# 12.1.1 Prerequisites
# 12.2 Tidy data
```{r}
table1
# Compute rate per 10,000
table1 %>% 
  mutate(rate = cases / population * 10000)

# Compute cases per year
table1 %>% 
  count(year, wt = cases)
# Visualise changes over time
library(ggplot2)
ggplot(table1, aes(year, cases)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))
```
# 12.2.1 Exercises
```{r error=TRUE}
#1. Using prose, describe how the variables and observations are organised in each of the sample tables.
?prose
#2. Compute the rate for table2, and table4a + table4b. You will need to perform four operations:
table2
##2.1. Extract the number of TB cases per country per year.
table2.cases<-table2  %>% filter(type=="cases") 
##2.2. Extract the matching population per country per year.
table2.pop<-table2  %>% filter(type=="population") 
##2.3. Divide cases by population, and multiply by 10000.
rate1<-table2.cases[,"count"]/table2.pop[,"count"]*10000
##2.4. Store back in the appropriate place.
#table2.pop$rate<-rate1 #
table2.cases[,"rate"]<-rate1 # How to add column in tidy way?
table2.cases
# how to use muatate()?
mutate(table2.cases,rate2=table2.cases[,"count"]/table2.pop[,"count"]*10000) # Error: not compatible with STRSXP
mutate(table2.cases,rate2=as.vector(table2.cases[,"count"]/table2.pop[,"count"]*10000)) # Error: not compatible with STRSXP ???

## Which representation is easiest to work with? Which is hardest? Why?
# table2 than table4 because there is only one year column, which enable to extract specific year.


#3. Recreate the plot showing change in cases over time using table2 instead of table1. What do you need to do first?
library(ggplot2)
ggplot(table2.cases, aes(factor(year), count)) + 
  geom_line(aes(group = country), colour = "grey50") + 
  geom_point(aes(colour = country))

```
# 12.3 Spreading and gathering
# 12.3.1 Gathering
```{r}
table4a
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases") # easy to use
table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
# To combine the tidied versions of table4a and table4b into a single tibble, we need to use dplyr::left_join(), which you’ll learn about in relational data.
tidy4a <- table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")
tidy4b <- table4b %>% 
  gather(`1999`, `2000`, key = "year", value = "population")
left_join(tidy4a, tidy4b)
```
# 12.3.2 Spreading
```{r}
# Spreading is the opposite of gathering. You use it when an observation is scattered across multiple rows. For example, take table2: an observation is a country in a year, but each observation is spread across two rows.
table2
spread(table2, key = type, value = count)

```
# 12.3.3 Exercises
#1. Why are gather() and spread() not perfectly symmetrical?
# Carefully consider the following example:
```{r}
stocks <- tibble(
  year   = c(2015, 2015, 2016, 2016),
  half  = c(   1,    2,     1,    2),
  return = c(1.88, 0.59, 0.92, 0.17)
)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`)
# (Hint: look at the variable types and think about column names.)
# 
# Both spread() and gather() have a convert argument. What does it do?
stocks %>% 
  spread(year, return)
stocks %>% 
  spread(year, return) %>% 
  gather("year", "return", `2015`:`2016`,convert=TRUE) # convet <chr> into <int> in year

```
#2. Why does this code fail?
```{r error=TRUE}
table4a %>% 
  gather(1999, 2000, key = "year", value = "cases")
#> Error in combine_vars(vars, ind_list): Position must be between 0 and n
table4a
# 1999 and 2000 are treated as value in integer. Treat them as characters.
table4a %>% 
  gather(`1999`, `2000`, key = "year", value = "cases")

```
#3. Why does spreading this tibble fail? How could you add a new column to fix the problem?
```{r error=TRUE}
people <- tribble(
  ~name,             ~key,    ~value,
  #-----------------|--------|------
  "Phillip Woods",   "age",       45,
  "Phillip Woods",   "height",   186,
  "Phillip Woods",   "age",       50,
  "Jessica Cordero", "age",       37,
  "Jessica Cordero", "height",   156
)
people
# trying spread
spread(people,key,value) # Error: Duplicate identifiers for rows (1, 3)
# adding new column with genertion
people %>% 
  mutate(generation=c("1st","1st","2nd","1st","1st")) %>%
  spread(key,value)

```
#4. Tidy the simple tibble below. Do you need to spread or gather it? What are the variables?
```{r error=TRUE}
preg <- tribble(
  ~pregnant, ~male, ~female,
  "yes",     NA,    10,
  "no",      20,    12
)
preg %>%
  gather(male,female,key=sex) # Error: Each variable must have a unique name.

is.na(preg) <-0
preg %>% 
  replace_na(list(male=0)) %>%
  gather(male,female,key="sex") # does not work

```
# 12.4 Separating and uniting
# 12.4.1 Separate
```{r}
table3
table3 %>% 
  separate(rate, into = c("cases", "population"))
table3 %>% 
  separate(rate, into = c("cases", "population"), sep = "/")

table3 %>% 
  separate(rate, into = c("cases", "population"), convert = TRUE)
table3 %>% 
  separate(year, into = c("century", "year"), sep = 2)
```
# 12.4.2 Unite
```{r}
table5 %>% 
  unite(new, century, year) # what happened to "paste()"?
table5 %>% 
  unite(new, century, year, sep = "")

```
# 12.4.3 Exercises
```{r error=TRUE}
#1. What do the extra and fill arguments do in separate()? Experiment with the various options for the following two toy datasets.
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"))
# test extra="drop"
tibble(x = c("a,b,c", "d,e,f,g", "h,i,j")) %>% 
  separate(x, c("one", "two", "three"),extra="drop")
# 
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"))
# test fill="right"
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"),fill="right")
# test fill="left"
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"),fill="left")

#2. Both unite() and separate() have a remove argument. What does it do? Why would you set it to FALSE?
table5
table5 %>% 
  unite(new, century, year) # what happened to "paste()"?
?unite
table5 %>% 
  unite(new, century, year,remove=FALSE) # keep original columns
# 
tibble(x = c("a,b,c", "d,e", "f,g,i")) %>% 
  separate(x, c("one", "two", "three"),remove=FALSE) # # keep original columns

#3. Compare and contrast separate() and extract(). Why are there three variations of separation (by position, by separator, and with groups), but only one unite?
?separate
# Given either regular expression or a vector of character positions, separate() turns a single character column into multiple columns.
df <- data.frame(x = c(NA, "a.b", "a.d", "b.c"))
df %>% extract(x, c("A", "B"),"([[:alnum:]]+).([[:alnum:]]+)") # does work
df %>% separate(x, c("A", "B"),"([[:alnum:]]+).([[:alnum:]]+)") # does not work
df %>% separate(x, c("A", "B")) # does work
df %>% separate(x, c("A", "B"),sep=".") # does not work. why?
df %>% separate(x, c("A", "B"),sep=".",extra="merge") # does not work. why?

df %>% separate(x, c("A", "B"),sep=.) # does not work

df %>% separate(x, c("A", "B"),c(1,2)) # by position? does work? No. How to do?

# If every row doesn't split into the same number of pieces, use
# the extra and file arguments to control what happens
df <- data.frame(x = c("a", "a b", "a b c", NA))
df %>% separate(x, c("a", "b"))
# The same behaviour but no warnings
df %>% separate(x, c("a", "b"), extra = "drop", fill = "right")
# Another option:
df %>% separate(x, c("a", "b"), extra = "merge", fill = "left")

# If only want to split specified number of times use extra = "merge"
df <- data.frame(x = c("x: 123", "y: error: 7"))
df %>% separate(x, c("key", "value"), ": ", extra = "merge")

?extract
# Given a regular expression with capturing groups, extract() turns each group into a new column. If the groups don't match, or the input is NA, the output will be NA.

df <- data.frame(x = c(NA, "a-b", "a-d", "b-c", "d-e"))
df %>% extract(x, "A")
df %>% extract(x, c("A", "B"), "([[:alnum:]]+)-([[:alnum:]]+)") # does  work
df %>% separate(x, c("A", "B"), sep="([[:alnum:]]+)-([[:alnum:]]+)") # does not work
df %>% separate(x, c("A", "B"), sep="-") # does work

?unite # Convenience function to paste together multiple columns into one.



```
